---
title: "CS 422 HW5"
author: "Sesha Shai Datta Kolli"
output:
  html_notebook:
    toc: yes
    toc_float: yes
  html_document:
    toc: yes
    df_print: paged
---


### Part 2.1-B - Frequent Itemsets and Rules of 1K transactions
```{r, fig.width = 10, fig.width = 10}
#install.packages("arules")
library(arules)
# 1k Transaction dataset
transactions_1k<-read.transactions(file = "tr-1k-canonical.csv",sep=",")
summary(transactions_1k)
inspect(transactions_1k[1:5])
itemFrequencyPlot(transactions_1k,support=0.01)

frequent_1k <- apriori(transactions_1k, parameter = list(supp = 0.01, conf = 0.8,target="frequent itemsets"))
#inspect(sort(frequent_1k,decreasing=TRUE, by="count"))
frequent_1k_df<-DATAFRAME(frequent_1k)
colnames(frequent_1k_df)<- c('ITEMS','Support')
print(frequent_1k_df[order(frequent_1k_df$Support, decreasing = TRUE), 1:2])


rls_1k <- apriori(transactions_1k, parameter = list(supp = 0.01, conf = 0.8,target="rules"))
rls_1k_df<-DATAFRAME(rls_1k)
colnames(rls_1k_df)<- c('Antecedent','Consequent','Support','Confidence')
print(rls_1k_df[order(rls_1k_df$Support, decreasing = TRUE), 1:4])
```



### Part 2.1-B- Frequent Itemsets and Rules of 5K transactions
```{r}

# 5K Transactions dataset
transactions_5k<-read.transactions(file = "tr-5k-canonical.csv",sep=",")

frequent_5k <- apriori(transactions_5k, parameter = list(supp = 0.01, conf = 0.8,target="frequent itemsets"))
#inspect(sort(frequent_5k,decreasing=TRUE, by="count"))
frequent_5k_df<-DATAFRAME(frequent_5k)
colnames(frequent_5k_df)<- c('ITEMS','Support')
print(frequent_5k_df[order(frequent_5k_df$Support, decreasing = TRUE), 1:2])


rls_5k <- apriori(transactions_5k, parameter = list(supp = 0.01, conf = 0.8,target="rules"))
rls_5k_df<-DATAFRAME(rls_5k)
colnames(rls_5k_df)<- c('Antecedent','Consequent','Support','Confidence')
print(rls_5k_df[order(rls_5k_df$Support, decreasing = TRUE), 1:4])
```

### Part 2.1-B- Frequent Itemsets and Rules of 20K transactions
```{r}
# 20K Transactions dataset
transactions_20k<-read.transactions(file = "tr-20k-canonical.csv",sep=",")

frequent_20k <- apriori(transactions_20k, parameter = list(supp = 0.01, conf = 0.8,target="frequent itemsets"))
#inspect(sort(frequent_20k,decreasing=TRUE, by="count"))
frequent_20k_df<-DATAFRAME(frequent_20k)
colnames(frequent_20k_df)<- c('ITEMS','Support')
print(frequent_20k_df[order(frequent_20k_df$Support, decreasing = TRUE), 1:2])


rls_20k <- apriori(transactions_20k, parameter = list(supp = 0.01, conf = 0.8,target="rules"))
rls_20k_df<-DATAFRAME(rls_20k)
colnames(rls_20k_df)<- c('Antecedent','Consequent','Support','Confidence')
print(rls_20k_df[order(rls_20k_df$Support, decreasing = TRUE), 1:4])
```

### Part 2.1-B- Frequent Itemsets and Rules of 75K transactions
```{r}
# 75K Transactions dataset
transactions_75k<-read.transactions(file = "tr-75k-canonical.csv",sep=",")

frequent_75k <- apriori(transactions_75k, parameter = list(supp = 0.01, conf = 0.8,target="frequent itemsets"))
#inspect(sort(frequent_75k,decreasing=TRUE, by="count"))
frequent_75k_df<-DATAFRAME(frequent_75k)
colnames(frequent_75k_df)<- c('ITEMS','Support')
print(frequent_75k_df[order(frequent_75k_df$Support, decreasing = TRUE), 1:2])

rls_75k <- apriori(transactions_75k, parameter = list(supp = 0.01, conf = 0.8,target="rules"))
rls_75k_df<-DATAFRAME(rls_75k)
colnames(rls_75k_df)<- c('Antecedent','Consequent','Support','Confidence')
print(rls_75k_df[order(rls_75k_df$Support, decreasing = TRUE), 1:4])
```
Note: Please maximize the window to view the output.

### Part 2.1-C


When we compare rules in 1000, 5000, 20000, and 75000 datasets, there is no difference in terms of support because there is only a 0.01 variation in maximum support in each dataset. The maximum support achieved in each dataset is roughly 0.040. If support = 0.01 and confidence = 0.8, all datasets produce roughly 80 rules.

The distribution of itemsets in the dataset changes as the number of transactions grows. Lift is steadily decreasing as well. The most common itemset in each dataset changes as the number of observations grows.


### Part 2.1-D-i
```{r}
#Most Frequent item or itemset
frequent_75k_df2<-DATAFRAME(frequent_75k)
colnames(frequent_75k_df2)<- c('ITEMS','Support','Count')
print(head(frequent_75k_df2[order(frequent_75k_df2$Count, decreasing = TRUE), 1:3],1))
```
### Part 2.1-D-ii
```{r}

#Least Frequent item or itemset
frequent_75k_df2<-DATAFRAME(frequent_75k)
colnames(frequent_75k_df2)<- c('ITEMS','Support','Count')
print(head(frequent_75k_df2[order(frequent_75k_df2$Count, decreasing = FALSE), 1:3],1))
```


